---
title: "Supplementary materials"
output: html_notebook
---

# 1. Load necessary materials 


## 1.1. Load necessary libraries
```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Load libraries for data manipulation, analysis, and visualization

library(readr)       # For reading CSV files efficiently
library(dplyr)       # For data manipulation (filter, mutate, summarize, etc.)
library(tidyr)       # For tidying data (pivoting, reshaping)
library(ggplot2)     # For creating high-quality plots
library(lme4)        # For fitting linear mixed-effects models
library(emmeans)     # For computing estimated marginal means (post-hoc comparisons)
library(lmerTest)    # For hypothesis tests in linear mixed-effects models
library(gridExtra)   # For arranging multiple plots in a grid layout
library(cowplot)     # For complex ggplot2 layouts and annotations
library(DescTools)   # For descriptive statistics and other statistical tools
library(sjmisc)      # For data preparation and variable recoding
library(readxl)      # For reading Excel files
library(sjPlot)      # For creating summary tables and plots (especially for mixed models)
library(broom.mixed) # For tidying mixed-model outputs into data frames
library(ordinal)     # For ordinal regression models
library(ggridges)    # For ridgelines
```

## 1.2. Load project datasets
```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Set working directory to the project folder
setwd(" ")

# Load CSV files into R and assign them to variables
# Each dataset is read using read_csv from the readr package

ai_prediction <- read_csv("ai_prediction.csv")       # Is it AI or not
funny_rating <- read_csv("funny_rating.csv")         # Human ratings of funniness
familiarity <- read_csv("familiarity.csv")           # Familiarity ratings
seen_before <- read_csv("seen_before.csv")           # Whether participants had seen items before
authors_gold_standard <- read_csv("authors_gold_standard.csv") # Gold standard labels 
```


# 2. Prepare and reshape datasets

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Add a 'subject' column as a unique identifier for each row
# Remove the first column (assumed to be an old index or redundant)
ai_prediction <- ai_prediction %>% mutate(subject = row_number()) %>% select(-1)
funny_rating  <- funny_rating %>% mutate(subject = row_number()) %>% select(-1)
familiarity   <- familiarity %>% mutate(subject = row_number()) %>% select(-1)
seen_before   <- seen_before %>% mutate(subject = row_number()) %>% select(-1)
authors_gold_standard <- authors_gold_standard %>% select(-1) # No subject needed

# 'pivot_longer' reshapes the data: each item becomes a row per subject
ai_long <- ai_prediction %>%
  pivot_longer(
    -subject,            # keep 'subject' column as is
    names_to = "item",    # column name for original variable names
    values_to = "ai_prediction"  # column name for values
  )

funny_long <- funny_rating %>%
  pivot_longer(
    -subject,
    names_to = "item",
    values_to = "funny_rating"
  )

familiarity_long <- familiarity %>%
  pivot_longer(
    -subject,
    names_to = "item",
    values_to = "familiarity"
  )

seen_before_long <- seen_before %>%
  pivot_longer(
    -subject,
    names_to = "item",
    values_to = "seen_before"
  )

authors_gold_standard_long <- pivot_longer(
  authors_gold_standard,
  cols = everything(),                    # pivot all columns
  names_to = "item",                      # column to hold original column names
  values_to = "authors_gold_standard"     # column to hold the values
)
```

# 3. Combine all datasets into a single long-format table

```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Merge all long-format datasets by 'subject' and 'item'
# Authors gold standard only by 'item' since it's the same for all subjects
data_long <- ai_long %>%
  left_join(funny_long, by = c("subject", "item")) %>%
  left_join(familiarity_long, by = c("subject", "item")) %>%
  left_join(seen_before_long, by = c("subject", "item")) %>%
  left_join(authors_gold_standard_long, by = "item") %>%
  mutate(
    # Convert relevant variables to factors for modeling
    ai_prediction = as.factor(ai_prediction),
    familiarity   = as.factor(familiarity),
    seen_before   = as.factor(seen_before),
    authors_gold_standard = as.factor(authors_gold_standard),
    
    # Ensure numeric variables are correctly typed
    funny_rating  = as.numeric(funny_rating),
    
    # Clean up 'item' names and convert to factor
    item          = as.factor(gsub("item", "", item)),
    
    # Convert subject ID to factor
    subject       = as.factor(subject)
  )
```


# 4. Relevel ordinal variables
```{r, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Convert selected variables to ordered factors
# This is important for ordinal modeling and analyses
data_final <- data_long %>%
  mutate(
    seen_before  = as.ordered(seen_before),   # 'seen_before' is now an ordered factor
    funny_rating = as.ordered(funny_rating)   # 'funny_rating' is now an ordered factor
  )
```



#5.  Statistical analysis

## Null model

```{r}
# Fit a null cumulative link mixed model (CLMM)
# This model includes only random intercepts for subject and item
# No fixed effects are included (intercept-only model)
m.null <- clmm(
  funny_rating ~ 1 + (1 | subject) + (1 | item),  # Formula for null model
  data = data_final
)

# Display model summary in the console (can be suppressed if needed)
summary(m.null)
```

## Model 1: Main effect of AI prediction
```{r}
# Fit a CLMM including AI prediction as a fixed effect
m1.ai.pred <- clmm(
  funny_rating ~ ai_prediction + (1 | subject) + (1 | item),
  data = data_final
)

# Show summary with coefficients and random effects
summary(m1.ai.pred)

# Likelihood ratio test comparing null model to Model 1
print(anova(m.null, m1.ai.pred))
```

## Model 2: Random slope of AI prediction by subject
```{r}
# Fit a CLMM with a random slope for AI prediction within subjects
# - This allows each subject to have their own effect of AI prediction
# - Items still have only random intercepts
m2.ai.pred <- clmm(
  funny_rating ~ ai_prediction + (1 + ai_prediction | subject) + (1 | item),
  data = data_final
)

# Show summary with coefficients, thresholds, and random effects
summary(m2.ai.pred)

# Likelihood ratio test: compare Model 1 (fixed effect only) vs. Model 2 (random slope)
anova(m1.ai.pred, m2.ai.pred)
```
Result:

The likelihood ratio test shows no significant improvement when adding a random slope for AI prediction.

Therefore, Model 1 (main effect of AI prediction only) is preferred.


## Model 3: Main effect of  gold standard
```{r}
# Fit a CLMM with AI prediction and authors' gold standard as fixed effects
m3.auth.std <- clmm(
  funny_rating ~ ai_prediction + authors_gold_standard + 
    (1 | subject) + (1 | item),
  data = data_final
)

# Show summary of Model 3
summary(m3.auth.std)

# Likelihood ratio test: compare Model 1 vs. Model 3
anova(m1.ai.pred, m3.auth.std)
```
Result:

Adding the gold standard as a fixed effect does not significantly improve model fit.

Model 1 (main effect of AI prediction only) remains the preferred model.

## Model 4: Main effect of familiarity
```{r}
# Fit a CLMM with AI prediction and familiarity as fixed effects
m4.fam <- clmm(
  funny_rating ~ ai_prediction + familiarity + 
    (1 | subject) + (1 | item),
  data = data_final
)

# Show summary of Model 4
summary(m4.fam)

# Likelihood ratio test: compare Model 1 (AI prediction only) vs. Model 4
anova(m1.ai.pred, m4.fam)
```
Result:

Adding familiarity as a fixed effect does not significantly improve the model fit.

Model 1 (main effect of AI prediction only) remains the preferred model.

## Model 5: Main effect of “seen before”
```{r}
# Fit a CLMM with AI prediction and 'seen_before' as fixed effects
m5.seen <- clmm(
  funny_rating ~ ai_prediction + seen_before + 
    (1 | subject) + (1 | item),
  data = data_final
)

# Show summary of Model 5
summary(m5.seen)

# Likelihood ratio test: compare Model 1 (AI prediction only) vs. Model 5
anova(m1.ai.pred, m5.seen)
```
Result:

Adding seen_before as a fixed effect significantly improves the model fit compared to Model 1.

At this stage, Model 5 is preferred over previous models.

## Model 6: Random slope of “seen before” by subject
```{r}
# Fit a CLMM with AI prediction and 'seen_before' as fixed effects
# Add a random slope for 'seen_before' within subjects
m6.seen <- clmm(
  funny_rating ~ ai_prediction + seen_before + 
    (1 + seen_before | subject) + (1 | item),
  data = data_final
)

# Show summary of Model 6
summary(m6.seen)

# Likelihood ratio test: compare Model 5 (fixed effect only) vs. Model 6 (random slope)
anova(m5.seen, m6.seen)
```
Result:

Adding a random slope for seen_before by subject does not significantly improve model fit compared to Model 5.

Therefore, Model 5 (main effect of AI prediction + seen_before) remains the preferred model.


## Model 7: Interaction between AI prediction and “seen before”
```{r}
# Fit a CLMM with an interaction between AI prediction and 'seen_before'
m7.inter <- clmm(
  funny_rating ~ ai_prediction * seen_before + 
    (1 | subject) + (1 | item),
  data = data_final
)

# Show summary of Model 7
summary(m7.inter)

# Likelihood ratio test: compare Model 5 (main effects only) vs. Model 7 (interaction added)
anova(m5.seen, m7.inter)
```
Result:

Adding the interaction between AI prediction and seen_before does not significantly improve model fit.

Therefore, the simpler Model 5 (main effects only) remains the preferred model.

## Final Model
```{r}
final_model <- m5.seen
summary(final_model)
```


Main results:

- When roast judged as written by human, it is rated as sigingicantly more funny
- the more the participants is sure the to have seen the roast before, the more it is funny

# 6. Plot results
```{r}

library(patchwork)
data_scaled$authors_gold_standard <- as.character(data_scaled$authors_gold_standard)
data_scaled$authors_gold_standard[data_final$authors_gold_standard == "human"] <- "Human"
# Example: suppose your predictors are ai_prediction, human_prediction, and seen_before
# (replace or add more variables as needed)

p1 <- ggplot(data_scaled, aes(x = ai_prediction, y = as.numeric(funny_rating_scaled))) +
  geom_boxplot(trim = FALSE, fill = "lightblue", color = "blue", outlier.shape = NA) +
  labs(
    subtitle = "Author prediction",
    x = "",
    y = "Funniness Rating\n(normalized per participant)"
  ) +
  theme_minimal() 

p2 <- ggplot(data_scaled, aes(x = authors_gold_standard, y = as.numeric(funny_rating_scaled))) +
  geom_boxplot(trim = FALSE, fill = "palegreen", color = "limegreen",outlier.shape = NA) +
  labs(
    subtitle = "Author",
    x="",
    y=""
  ) +
  theme_minimal()

p3 <- ggplot(data_scaled, aes(x = seen_before, y = as.numeric(funny_rating_scaled))) +
  geom_boxplot(trim = FALSE, fill = "lightpink", color = "magenta",outlier.shape = NA) +
  labs(
    subtitle = "Prior encounter",
    x="",
    y = "Funniness Rating\n(normalized per participant)"
  ) +
  theme_minimal()

p4 <- ggplot(data_scaled, aes(x = familiarity, y = as.numeric(funny_rating_scaled))) +
  geom_boxplot(trim = FALSE, fill = "purple", color = "purple4",outlier.shape = NA) +
  labs(
    subtitle = "Familiarity",
    x="",
    y=""
  ) +
  theme_minimal()

p1 <- p1 + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p2 <- p2 + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p3 <- p3 + theme(axis.text.x = element_text(angle = 45, hjust = 1))
p4 <- p4 + theme(axis.text.x = element_text(angle = 45, hjust = 1))

p2 <- p2 + ylab(NULL)
#p3 <- p3 + ylab(NULL)
p4 <- p4 + ylab(NULL)

# Combine them into one layout using patchwork
combined_plot <- (p1 | p2) / (p3 | p4) +
  plot_layout(guides = "collect") 


combined_plot

# Save as a square image (e.g., 8x8 inches)
ggsave(
  filename = "combined_plot.png",   # change to your desired file name
  plot = combined_plot,             # your patchwork plot
  width = 4,                        # width in inches
  height = 6,                       # height in inches (same as width for square)
  dpi = 300                         # resolution
)
```


